Spatial Reasoning and Statistical Graphics

### 2014-11-06 Paper/Draft1/TestingVisualAptitude.Rnw
some re-phrasing

\subsection{Linear model of demographic factors}
\newtext{The demographic variables are highy correlated, e.g.~there is a high correlation between STEM majors and taking calculus 1 -- only eight students are either not STEM, but have taken calculaus 1 or are STEM but have not taken the class. The situation between a statistics class and doing research is even more correlated. Only one student is doing research who has (not yet ?) taken a statistics course. A principal component of the five math/stats questions splits the variables into two areas: the first principal component is an average of math skills, calculus 1 and STEM, while the second principal component is an average of having taken a statistics class and doing research. We therefore decided to use sums of these variables to come up with a separate math and a stats score. Note, that the correlation between the math and the stats score is almost zero. }
<<linearmodel, results='hide', echo=FALSE>>=
pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
ans.model <- data.frame(ans.summary, data.frame(pca2$x))
# exclude all variables that are not supposed to go into the linear model
# those are the unrotated tests; instead we have the PCs
excl <- c("id", "card_rot", "fig_class", "folding", "vis_search", "vidgame_hrs", "major1", "major2", "minor1", "minor2", "learning_disability", "colorblind", "epilepsy", "normal_vision")

@@

# use AIC in backward selection
library(MASS)
m1 <- stepAIC(m0, direction="both")
@
\newtext{We fit a linear model of lineup scores in all of the demographic variables and the test scores from the visuo-spatial tests. Then we use a step-wise backwards selection based on AIC (Akaike Information Criterion - do we need a reference?) to select the best model. The result is shown in table~\ref{tab:m1}. Only two covariates stay in the model: PC1 and MATH, reflecting two dimensions of what affects lineup scores. We can think of PC1 as an innate visual ability, while the MATH score is a matter of both ability and training. 
}
<<modelresults, echo=FALSE, results='asis'>>=
library(xtable)
print(xtable(summary(m1), caption="Estimates and significances of a linear model of lineup scores.", label="tab:m1"), caption.placement="top")
@
\subsection{Lineup Types}
\comment{
results need to go here
}

>>>>

\subsection{Linear model of demographic factors}
\newtext{Note that all of the demographic variables in the survey are highy correlated, for example there is a high correlation between STEM majors and taking calculus.% -- only eight students are either not STEM, but have taken calculaus 1 or are STEM but have not taken the class. 
Similaryl, the correlation  between having taken a statistics class and having been involved in mathematics or statistics research is high. Only one student is doing research who has (not yet ?) taken a statistics course. A principal component of the five math/stats questions splits the variables into two main areas: the first principal component is an average of math skills, calculus 1 and STEM, while the second principal component is an average of having taken a statistics class and doing research. We therefore decided to use sums of these variables to come up with a separate math and a stats score. Note, that the correlation between the math and the stats score is almost zero. }
<<linearmodel, results='hide', echo=FALSE>>=
pca2 <- prcomp(ans.summary[,c("card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
ans.model <- data.frame(ans.summary, data.frame(pca2$x))
# exclude all variables that are not supposed to go into the linear model
# those are the unrotated tests; instead we have the PCs
excl <- c("id", "card_rot", "fig_class", "folding", "vis_search", "vidgame_hrs", "major1", "major2", "minor1", "minor2", "learning_disability", "colorblind", "epilepsy", "normal_vision")

@@

# use AIC in backward selection
library(MASS)
m1 <- stepAIC(m0, direction="both")
@
\newtext{We fit a linear model of lineup scores in the thus modified demographic variables and the test scores from the visuo-spatial tests. Using a step-wise backwards selection based on AIC (Akaike Information Criterion - do we need a reference?) we select the best model. The result is shown in table~\ref{tab:m1}. Only two covariates stay in the model: PC1 and MATH, reflecting two dimensions of what affects lineup scores. We can think of PC1 as an innate visual ability, while the MATH score is a matter of both ability and training. 
}
<<modelresults, echo=FALSE, results='asis'>>=
library(xtable)
print(xtable(summary(m1), caption="Estimates and significances of a linear model of lineup scores.", label="tab:m1"), caption.placement="top")
@
\subsection{Lineup Types}
\comment{
results need to go here
}
